---
title: "Visualize EBV Detection"
author: "Veronica Russell, Dr. Dave Lab Rotation"
date: "5/6/2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

#Description:
##INPUT: 
Script takes as input a user constructed .csv file titled "star_ebvmasked_count_f4F1024.csv" with the following column headers [required headers are sample_id, analysis_id, ebv_positive, sample_type, mean_coverage, ebv_count, and unique_ebv_count]
It is recommended to have more than 5 samples per panel


__________________________________________________
####required column headers:
sample_id - davelab Sample Id of form #_T_1

analysis_id - davelab Analysis ID number

panel - categorical "1MB", "8MB", "110MB", or "other"

ebv_positive - categorical "yes", "weak", "no", "unknown" known EBV positive sample, determined from WHO pathology reports at https://who.davelab.org

sample_type - categorical "FFPE","cell line", "other"

mean_coverage - RNA sample mean coverage determined from qc report at db.davelab.org under depth_of_coverage_aligned_read_stats

ebv_count - ebv reads (single) count determined from sum of quantified reads of ebv_detection module output ebv_detection_EBV_detection.{SAMPLE_ID}_A{ANALYSIS_ID}_ebv_paired_Aligned.out.sam and ebv_detection_EBV_detection.{SAMPLE_ID}_A{ANALYSIS_ID}_ebv_singles_Aligned.out.sam

unique_ebv_count - unique read (single) counts determined from sum of unique reads extracted from ebv_detection module output ebv_detection_EBV_detection.{SAMPLE_ID}_A{ANALYSIS_ID}ebv_paired_Log.final.out and ebv_detection_EBV_detection.{SAMPLE_ID}_A{ANALYSIS_ID}ebv_singles_Log.final.out



####optional headers 
*can be used as alternates for mean_coverage at all locations in this script (may require information outside of pathology report, qc report, or ebv_detection analysis):

total_sample_readprs - from qc report at db.davelab.org under Total TNA stats Total Input Read Pairs

total_rna_readprs - from qc report at db.davelab.org under Total TNA stats Total RNA Read Pairs

total_human_unmapped_reads - processed reads reported by samtools view -f4 -F1024 {IN_BAM} -c divided by 2 (for pairs)

mapped_pls_unmapped_reads - processed reads reported by samtools view -F3328 {IN_BAM} -c divided by 2 (for pairs) 

__________________________________________________

##OUTPUT
1) Visualization summary plot of full .csv file and all panels
2) Visualization of individual panels


#Setup:
##Knitr Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load Libaries
```{r}
library(ggplot2)
library(tidyverse)
```

##Set themes (provided by Dr. Sara Selitsky)
```{r}
theme_sara_90 <- function() {
  theme_bw(base_size=14)+
    theme(axis.text.x=element_text(angle=90,hjust = 1,vjust = 0.5),
          axis.text=element_text(color="black"),
          panel.background=element_rect(color="black"),
          strip.text = element_text(size=12),
          strip.background = element_rect(fill="white"))
}
```


##Import data
```{r}
star_ebv_data <- read.csv(file='star_counts_ebvmasked_f4F1024.csv')
head(star_ebv_data)
```


#Data Processing:
```{r}
#fix where zero with adding 1/2 smallest zero value to all zeroes then taking log
star_ebv_data$ebv_count[star_ebv_data$ebv_count==0]<- 1
star_ebv_data$log_ebv_count <- log(star_ebv_data$ebv_count)
head(star_ebv_data)

#Split data for panel types included in dataframe
by_panel <- split(star_ebv_data, star_ebv_data$panel)

#select column variable indicative of graph point size (default=mean_coverage)
#if using optional column headers can replace "mean_coverage" with any desired measure of sample total reads or quality
size_var=star_ebv_data$mean_coverage
```

#Graphs:
##Metadata Sample Coverage Boxplots
```{r}
ggplot(star_ebv_data, aes(x=panel, y=mean_coverage))+xlab("panel")+ylab("mean_coverage")+geom_boxplot()
```


##All Panel Summary Graph
```{r}
#if using optional column headers can replace "mean_coverage" with any desired measure of sample total reads or quality
gg_star_all_samples <- ggplot(star_ebv_data, aes(x=reorder(sample_id, log_ebv_count), y=log_ebv_count, color=ebv_positive, size=size_var)) + geom_point(aes(shape=panel), alpha=0.4) +   scale_color_manual(name = "Known EBV Status",
                     values = c("yes" = "green",
                                "weak" = "blue",
                                "no" = "red",
                                "unknown" = "gray"),
                     labels = c("yes" = "Positive",
                                "weak" = "Weak Positive",
                                "no" = "Negative",
                                "unknown" = "Unknown")) +
scale_shape_manual(name = "Davelab Panel",
                   values=c("1MB" = 17,
                            "8MB" = 15,
                            "110MB" = 16,
                            "other" = 18)) +
scale_size_continuous(name="Mean RNA Coverage") +
theme_sara_90() + labs(title="STAR Log(EBV Read Count) by Sample ID", x="Sample ID", y="Log(EBV Read Counts)")

gg_star_all_samples
```


##Individual Panels Graphs
```{r}

for (x in names(by_panel)) {
  panel <- subset(star_ebv_data, panel==x)
  #if using optional column headers can replace "mean_coverage" with any desired measure of sample total reads or quality
  size_var2 <- panel$mean_coverage
  plot_title=paste("gg_star_",x,"_panel",sep="")
  plot_title <- ggplot(panel, aes(x=reorder(sample_id, log_ebv_count), y=log_ebv_count, color=ebv_positive, size=size_var2)) + geom_point(aes(shape=panel), alpha=0.4) + 
     scale_color_manual(name = "Known EBV Status",
                     values = c("yes" = "green",
                                "weak" = "blue",
                                "no" = "red",
                                "unknown" = "gray"),
                     labels = c("yes" = "Positive",
                                "weak" = "Weak Positive",
                                "no" = "Negative",
                                "unknown" = "Unknown")) +
      scale_shape_manual(name = "Davelab Panel",
                     values=c("1MB" = 17,
                              "8MB" = 15,
                              "110MB" = 16,
                              "other" = 18)) +
     scale_size_continuous(name="Mean RNA Coverage") +
     theme_sara_90() + labs(title=paste(x,"STAR Log(EBV Read Count) by Sample ID",sep=" "), x="Sample ID", y="Log(EBV Read Counts)")

  print(plot_title)
}

```


